---
title: "Coursera Practical Machine Learning Final Project"
subtitle: "OverView"
---
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The data consists of a Training data and a Test data (to be used to validate the selected model).

The goal of your project is to predict the manner in which they did the exercise. This is the “classe” variable in the training set. You may use any of the other variables to predict with.

---
subtitle: "Data Loading and Processing"
---

```{r}

library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(corrplot)

```
---
Title : Getting, Cleaning and Exploring the data
---

```{r}
train_in <- read.csv('./pml-training.csv', header=T)
valid_in <- read.csv('./pml-testing.csv', header=T)
dim(train_in)

```
---
Cleaning the input data
---

```{r}
trainData<- train_in[, colSums(is.na(train_in)) == 0]
validData <- valid_in[, colSums(is.na(valid_in)) == 0]
dim(trainData)
dim(validData)
```
---
Filter Datasets
---

```{r}
trainData <- trainData[, -c(1:7)]
validData <- validData[, -c(1:7)]
dim(trainData)
dim(validData)
```

---
Preparing the datasets for prediction
---

```{r}
set.seed(1234) 
inTrain <- createDataPartition(trainData$classe, p = 0.7, list = FALSE)
trainData <- trainData[inTrain, ]
testData <- trainData[-inTrain, ]
dim(trainData)
dim(testData)
```
---
Cleaning even further by removing the variables that are near-zero-variance
---
```{r}
NZV <- nearZeroVar(trainData)
trainData <- trainData[, -NZV]
testData  <- testData[, -NZV]
dim(trainData)
dim(testData)

```
---

After this cleaning we are down now to 53 variables

The following correlation plot uses the following parameters (source:CRAN Package ‘corrplot’) “FPC”: the first principal component order. “AOE”: the angular order tl.cex Numeric, for the size of text label (variable names) tl.col The color of text label.
---
```{r}

cor_mat <- cor(trainData[, -53])
corrplot(cor_mat, order = "FPC", method = "color", type = "upper", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))

```

```{r}

highlyCorrelated = findCorrelation(cor_mat, cutoff=0.75)

```

```{r}

names(trainData)[highlyCorrelated]

```



Model building
For this project we will use two different algorithms, classification trees and random forests, to predict the outcome.

classification trees
random forests
Generalized Boosted Model

Prediction with classification trees :


```{r}

set.seed(12345)
decisionTreeMod1 <- rpart(classe ~ ., data=trainData, method="class")
fancyRpartPlot(decisionTreeMod1)


```




